What words ought to exist? \\ {\normalsize Coining with coinduction}

### 2011-03-12 wishlist/paper.tex
decided to use gnuplot instead of illustrator for graph; will undo cdf junkj

My website ``{\sf snoot.org}'' has a number of games on it, including
a Scrabble clone called Scribble\footnote{{\tt
http://snoot.org/toys/scribble/}} and Boggle clone called
Muddle.\!\footnote{{\tt http://snoot.org/toys/muddle/}} This website
has been running for almost ten years, comprising over 150,000
Scribble games totaling 3.8 million words placed and 628,000 Muddle
games with over 10 million words found. During each game, players
repeatedly attempt to play words that aren't real. The computer
rebukes them, but hope really springs eternal with these people. It's
like they truly deeply wish to break out of the shackles of the
Official Scrabble Players Dictionary. So the first approach to
determining what words ought to exist is to analyze the words that
people tried to play, in order to try to extract the essence of
word-yearning.

This analysis is quite straightforward. I took the ten years of logs
files and extracted each attempt to play a word in Scribble or Muddle.
These log files are quite large, so the first step is just to get a
count, for each alleged word, and store those in a more convenient
format. There were XXX total words attemped in Scrabble and XXX in
Muddle. The most frequent ones appear in
Figure~\ref{fig:mostfrequent}. The most frequent words are all
legitimate words, since players have a bias towards attempting words
that will not be rebuked by the computer. Seeing the words that people
wish existed is a simple matter of filtering out the words that
already exist, using the Scrabble dictionary. XXX\% of the words
attempted in Scribble and XXX\% in Muddle were not real. The most
frequent ones appear in Figure~\ref{fig:mostfrequentfake}.

\begin{figure}
\caption{Someone needs to fill this in with the data.}
\label{fig:mostfrequent}
\end{figure}

\begin{figure}
\caption{Someone needs to fill this in with the data.}
\label{fig:mostfrequentfake}
\end{figure}

XXX I need to write an analysis of the list and put it here.

In my opinion these are pretty good candidates for words to exist.
Probably if you were playing someone really intense in Scrabble, and
he or she played one of these, and was super deadpan about it and
maybe had caused some crying contests before, and a known
sesquipedalianist, you would let these fly because they look like real
words to me. A point in their favor is that they are quite low-scoring
words in Scrabble; one might expect that the words most wished for are
like ``qzkwv'' for the scoring potential. The effect is probably due
to a few factors: Players are less likely to attempt obvious
non-words, common letters appear more often on the rack and on the
board and so the opportunity to play words like in
Figure~\ref{fig:mostfrequentfake} presents itself more frequently, and
in Muddle, there is no advantage to using unusual letters, except the
joy of being a weirdo. Nonetheless, this list is surely biased by the
specifics of Scribble and Muddle, and the question at hand is not just
what words ought to exist for the purpose of internet word games, but
for general purposes.

@@

Another downside is that this method completely ignores the many words
that are attempted only once or a small number of times. The
distribution is heavy-tailed; XXX\% of the words attempted are not
even in the top XXX (Figure~\ref{fig:gamedistribution}). Though those
words individually are not good candidates to exist, like tiny stars
wished upon in the night sky,\!\footnote{Astronomers now agree that
stars do exist, by the way.} in aggregate they form a significant
planetarium that may tell us what {\it kind} of words people wish
existed. For example, if we saw that the words ``sweeeeeeet'',
``sweeeeeeeeeeeeet'', ``sweeeet'' and ``sweeeeeeeeeeeeeet'' occurred a
few times each, we could infer that people wished that words like
``sweet'' with strictly more than two {\it e}s were real words. They
might even be indifferent to the absolute number of {\it e}s, as long
as there existed some legal variation with more than two {\it e}s.
(This appears to be borne out by data. According to Google's
estimates, the words ``swe$^n$t'' for various medium-sized $n$
(10--20) appear on the Internet with similar frequency. The only
exception is ``sweeeeeeeeeeeeeeeeeeet'', with 19 {\it e}s, which
unexpectedly appears three times as often as 18 or 20 {\it e}s does.)
In order to lance these two boils, in the next section I explore
statistical methods for generalizing from lots of individual examples.

@@

@@

Rob: etsy, nuuog
Chris: tapping on the bisphenol-A, nurm
David: wafflucinations
Lea: hnfff sound

@@

>>>>

My website ``{\sf snoot.org}'' has a number of games on it, including
a Scrabble clone called Scribble\footnote{{\tt
    http://snoot.org/toys/scribble/}} and Boggle clone called
Muddle.\!\footnote{{\tt http://snoot.org/toys/muddle/}} This website
has been running for almost ten years, comprising over 150,000
Scribble games totaling 3.8 million words placed and 628,000 Muddle
games with over 10 million words found. During each game, players
repeatedly attempt to play words that aren't real. The computer
rebukes them, but hope really springs eternal with these people. It's
like they truly deeply wish to break out of the shackles of the
Official Scrabble Players Dictionary.\!\footnote{For the analyses in
  this paper that depend on a list of legal words, I actually use a
  modified version of SOWPODS, which is the tournament list used in
  Australia and the UK, and significantly more permissive than the US
  Tournament Word List. Though the modified version is non-canonical,
  I stuck with it because it's what's been in use on the site for
  ten years.} So the first approach to determining what
words ought to exist is to analyze the words that people tried to
play, in order to try to extract the essence of word-yearning.

This analysis is quite straightforward. I took the ten years of logs
files and extracted each attempt to play a word in Scribble or Muddle.
These log files are quite large, so the first step is just to get a
count, for each alleged word, and store those in a more convenient
format. There were 3,572,226 total words attempted\footnote{Here a
  word attempted is the major word of the play. This does not include
  incidental words (typically two-letter ones) formed in the
  perpendicular direction.} in Scribble and 13,727,511 in Muddle. The
most frequent ones appear in Figure~\ref{fig:mostfrequent}. Aside from
the one-letter ones, the most frequent words are legitimate words,
since players have a bias towards attempting words that will not be
rebuked by the computer.

Seeing the words that people wish existed is a simple matter of
filtering out the words that already exist, using the Scrabble
dictionary. (I also filtered out one-letter ``words''. It is easy to
see that no one-letter words should exist, again because of
ambiguities created in spelling bees. Not only when literally spelling
``bees'', but according to the official Scripps National Spelling Bee
rules, the speller may optionally pronounce the word to be spelled
before and after spelling it. So if ``s'' were a word, then the
following ridiculous exchange obtains: Judge: ``S. The letter {\it s}.
Etruscan origin.'' Speller: ``S. S. S.'' and the judge cannot tell if
the speller meant to state the word before and after, or thinks the
word is spelled ``sss''.) 22.3\% of the words attempted in Scribble
and 36.8\% in Muddle were not real. The most frequent ones appear in
Figure~\ref{fig:mostfrequentfake}.

\begin{figure}
\begin{tabular}{|rl@{\qquad}rl|}
\multicolumn{2}{l}{{\bf \large Scribble}} &
\multicolumn{2}{l}{{\bf \large Muddle}} \\
\hline
Count & Word & Count & Word \\
\hline
45,605  &  a        &     20,412  &  late  \\
42,315  &  i        &     19,405  &  rate  \\
32,499  &  d$^*$    &     19,276  &  dear  \\
12,981  &  in       &     19,049  &  tear  \\
12,851  &  oe       &     19,019  &  date  \\
12,528  &  s$^*$    &     18,771  &  lear  \\
12,207  &  re       &     18,423  &  deal  \\
11,159  &  tv       &     18,231  &  real  \\
10,720  &  jo       &     18,138  &  lead  \\
10,386  &  it       &     18,076  &  tale  \\
10,369  &  et       &     17,969  &  lane  \\
9,659   &  qua      &     17,956  &  sear  \\
9,218   &  xi       &     17,570  &  read  \\
9,099   &  go       &     17,193  &  teal  \\
9,052   &  ow       &     17,170  &  lean  \\
8,801   &  qat      &     17,071  &  dare  \\
8,602   &  aa       &     16,923  &  dale  \\
8,278   &  un       &     16,892  &  seal  \\
8,142   &  en       &     16,806  &  sale  \\
8,005   &  or       &     16,465  &  seat  \\
\hline
\end{tabular}
\caption{Most frequently attempted words in Scribble and Muddle. Asterisks
indicate non-words.}
\label{fig:mostfrequent}
\end{figure}

\begin{figure}

\begin{tabular}{|rl@{\qquad}rl|}
\multicolumn{2}{l}{{\bf \large Scribble}} &
\multicolumn{2}{l}{{\bf \large Muddle}} \\
\hline
Count & Word & Count & Word \\
\hline
11,159 &  tv         &     16,251 &  dane    \\
4,003  &  ok         &     6,156  &  rane    \\
2,862  &  iraq       &     5,603  &  sare    \\
2,725  &  zen        &     5,576  &  nate    \\
2,448  &  cho        &     4,863  &  mear    \\
1,538  &  viz        &     4,750  &  cale    \\
1,418  &  sdasda     &     4,616  &  nees    \\
1,396  &  von        &     4,568  &  nale    \\
1,136  &  etc        &     4,507  &  fale    \\
878    &  int        &     4,347  &  deat    \\
829    &  june       &     4,263  &  tean    \\
745    &  lp         &     4,251  &  nile    \\
719    &  zion       &     4,160  &  mens    \\
665    &  cia        &     4,087  &  deel    \\
661    &  jim        &     3,851  &  deam    \\
651    &  iraqi      &     3,828  &  dana    \\
648    &  ques       &     3,781  &  beed    \\
542    &  que        &     3,769  &  lans    \\
502    &  tim        &     3,725  &  tade    \\
\hline
\end{tabular}
\caption{Most frequently attempted non-words in Scrabble and Muddle.}
\label{fig:mostfrequentfake}
\end{figure}

There's a clear difference between these two lists. The Scribble list
is dominated by words involving difficult-to-play letters like {\it v}
(there are no legal two-letter {\it v}-words). Most of the words would
probably be acknowledged as real, just not legal in Scribble. The ones
that don't already have meanings, like ``cho'' and ``int'' and ``que''
seem to be pretty good candidates to exist. The Muddle list is all
four-letter words (the minimum allowed length) using common letters.
Other than the ones that are already words, like ``dane'' and ``nile''
and ``mens'' (as in ``mens section'' or ``the powerfuel weapon kills
hard so many mens''), these are all good candidates for words to
exist. Probably if you were playing someone really intense in
Scrabble, and he or she played one of these, and was super deadpan
about it and maybe had caused some crying contests before, and a known
sesquipedalianist, you would let these fly because they look like real
words to me. A point in their favor is that they would be quite
low-scoring words in Scrabble; not a {\it z} or {\it q} to be found.
Even in the Scribble list there's no ``qzkwv'' junk. The effect is
probably due to a few factors: Players are less likely to attempt
obvious non-words, common letters appear more often on the rack and on
the board and so the opportunity to play words like in
Figure~\ref{fig:mostfrequentfake} presents itself more frequently, and
in Muddle, there is no advantage to using unusual letters, except the
joy of being a weirdo. Nonetheless, these lists are surely biased by
the specifics of Scribble and Muddle, and the question at hand is not
just what words ought to exist for the purpose of internet word games,
but for general purposes.

@@

Another downside is that this method completely ignores the many words
that are attempted only once or a small number of times. The
distribution is heavy-tailed; of the 564,610 unique words attempted,
501,939 of them aren't real! XXX\% of the words attempted are not
even in the top XXX (Figure~\ref{fig:gamedistribution}). Though those
words individually are not good candidates to exist, like tiny stars
wished upon in the night sky,\!\footnote{Astronomers now agree that
stars do exist, by the way.} in aggregate they form a significant
planetarium that may tell us what {\it kind} of words people wish
existed. For example, if we saw that the words ``sweeeeeeet'',
``sweeeeeeeeeeeeet'', ``sweeeet'' and ``sweeeeeeeeeeeeeet'' occurred a
few times each, we could infer that people wished that words like
``sweet'' with strictly more than two {\it e}s were real words. They
might even be indifferent to the absolute number of {\it e}s, as long
as there existed some legal variation with more than two {\it e}s.
(This appears to be borne out by data. According to Google's
estimates, the words ``swe$^n$t'' for various medium-sized $n$
(10--20) appear on the Internet with similar frequency. The only
exception is ``sweeeeeeeeeeeeeeeeeeet'', with 19 {\it e}s, which
unexpectedly appears three times as often as 18 or 20 {\it e}s does.)
In order to lance these two boils, in the next section I explore
statistical methods for generalizing from lots of individual examples.

@@

The letters across the top are the ``previous letter'' and the ones
across the left are the ``next letter'' and the box contains the
corresponding count. For example, the pair ``el'' appears four times.
(Pairs of letters are called ``bigrams'' by nerds, some nerd-poseurs,
and Markov who I can't tell if he was a nerd by his picture, because
he does have a pretty austere beard, but also did a lot of math.) One
of the useful things about a Markov chain is that it lets us predict
the next letter that we might see. For example, if we see ``half'',
then the column labeled {\bf f} above tells us that the next letter is
twice as often an {\it e} than a {\it u}, and that no other letters
ever occurred. Typically we think of these as being probabilities
inferred from our observations, so we say there's a 2/3 chance of {\it
e} following {\it f} and a 1/3 chance of {\it u}. Now the word ``llllll''
isn't so likely any more, because there's only a 1/4 chance of the
next letter being {\it l} once we see {\it l}.

Words are not just their interiors; it's also important what letters
tend to start and end words. We can do this by imagining that each
word starts and ends with some fake letters, and include those in the
Markov chain. Let's use {\bf \<} for the start symbol and {\bf \>} for
the end. So we pretend we observed ``\<hello\>'', ``\<helpful\>'', and
``\<felafel\>''. Speaking of which, could you imagine if there were such
a thing as a helpful felafel? Would you eat it? Because then it
probably can't help you any more, except to get fat.

\vspace{0.2in}
\begin{centering}
\begin{tabular}{|c|r|r|r|r|r|r|r|r|r|} % {helopfua}
\hline
\,  &  {\bf \<}    & {\bf h}   & {\bf e}   & {\bf l}   & {\bf o}   & {\bf p}   & {\bf f}   & {\bf u}   & {\bf a}    \\
\hline  %  \<     h     e     l     o     p     f     u     a
{\bf h} &  2  &  0   & 0   & 0   & 0   & 0   & 0   & 0   & 0    \\
\hline    
{\bf e} &  0  &  2   & 0   & 0   & 0   & 0   & 2   & 0   & 0    \\
\hline    
{\bf l} &  0  &  0   & 4   & 1   & 0   & 0   & 0   & 1   & 0    \\
\hline    
{\bf o} &  0  &  0   & 0   & 1   & 0   & 0   & 0   & 0   & 0    \\
\hline    
{\bf p} &  0  &  0   & 0   & 1   & 0   & 0   & 0   & 0   & 0    \\
\hline    
{\bf f} &  1  &  0   & 0   & 0   & 0   & 1   & 0   & 0   & 1    \\
\hline    
{\bf u} &  0  &  0   & 0   & 0   & 0   & 0   & 1   & 0   & 0    \\
\hline    
{\bf a} &  0  &  0   & 0   & 1   & 0   & 0   & 0   & 0   & 0    \\
\hline    
{\bf \>} &  0  &  0   & 0   & 2   & 1   & 0   & 0   & 0   & 0    \\
\hline
\end{tabular}
\end{centering}
\vspace{0.2in}

We just added these like other letters, but since the beginning symbol
{\bf \<} never occurs after other letters, we don't need a row for it
(it would be all zeroes), and similarly since no letters ever follow
{\bf \>} we don't need a column for it. Now the word ``lllll'' is
impossible because no words start with {\it l}.

\begin{verbatim}
0.166666667 hel
0.166666667 helpful
0.166666667 hello
0.083333333 felpful
0.083333333 fello
0.083333333 fel
0.041666667 helafel
0.041666667 helafelpful
0.041666667 helafello
\end{verbatim}

% laplace smoothing?
% TODO: the empty string?

TODO: most likely words in italian, too.

@@

Rob: etsy, nuuog
Chris: tapping on the bisphenol-A, nurm
David: wafflucinations
Lea: hnfff sound
Reed: pansepticon
Jessica: gruntle (like not disgruntled)

@@

% TODO: backformation, stripping dis- and anti- and un- from words
% then checking to see if they're still words?

% ``Scrallbe'' is kind of like god mode for scrabble.
